---
title: 'ManyWebcams: web-based eye-tracking'
author: "ManyWebcams"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: inline
---

When loading the packages for this script, you can choose to use [groundhog](https://groundhogr.com/) by setting `use_groundhog = T`. Groundhog will install packages from `2022-10-09` and ensures that the functions keep working the same way as when the script was last developed. If you prefer to load packages the conventional way, leave `use_groundhog = F` (default).

```{r libraries, results='hide', message=FALSE, warning=FALSE}
# should R use groundhog
use_groundhog <- F

# set vector w/ packages
packages = c("readr",
             "knitr",
             "tidyr",
             "dplyr",
             "ggplot2",
             "psych",
             "effsize",
             "Matrix",
             "lme4",
             "lmerTest",
             "carData",
             "effects",
             "stringr",
             "here",
             "glmmTMB")

# load packages
if(use_groundhog == T){
  packages_date <- "2022-10-09" 
  install.packages("groundhog")
  library(groundhog) # v2.0.1
  groundhog::groundhog.library(pkg = packages, date = packages_date)
  rm(packages, packages_date, use_groundhog) # clean
}else{
  # if necessary install
  installed_packages <- packages %in% rownames(installed.packages())
  if(any(installed_packages == F)){install.packages(packages[!installed_packages])}
  # load packages
  invisible(lapply(packages, library, character.only = T))
  # clean
  rm(installed_packages, packages, use_groundhog)
}

knitr::opts_chunk$set(cache = TRUE, warn = FALSE, message = FALSE)
options(scipen=999)
```

# Intro
Analysis script for the ManyWebcams web-based eye-tracking study.

# Setting constants
```{r}
FILE_NAME_PILOT <- "pilot_data.csv"

FILE_NAME_WEBCAM <- "relative_data.csv"
FILE_NAME_DEMO <- "demographic_data.csv"
FILE_NAME_EXCLUSION_AFTERPRE <- "participants_excluded_after_pre.txt"

# already excluded from the datasets, read for reporting
FILE_NAME_EXCLUSION_MANUAL <- "excluded_trials.csv"
FILE_NAME_EXCLUSION_SAMPLING <- "excluded_trials_samplingrate.txt"
  
DIR_NAME_DATA_MERGE <- "lab_data"
DIR_NAME_DATA_MERGE_OUTPUT <- "merge_output"

FILE_NAME_SAMPLING_RATE <- "transformed_data.csv"
```

# File reading and processing
```{r}

data_pilot <- read.csv(FILE_NAME_PILOT, stringsAsFactors = TRUE)

# check if data merge is necessary
if(dir.exists(DIR_NAME_DATA_MERGE)){
  data_webcam_list <- list()
  data_demo_list <- list()
  
  posterior_exclusions <- c()
  samplingrate_exclusions <- c()
  manual_exclusions_list <- list()
  
  # at the moment only reproducible with the lab_data folder in the repo
  sampling_rate_data_list <- list()
  
  labs <- list.files(path=DIR_NAME_DATA_MERGE)
  for(lab in labs){
    
    data_webcam_part <- read.csv(file.path(DIR_NAME_DATA_MERGE, lab, FILE_NAME_WEBCAM), stringsAsFactors = TRUE)
    data_webcam_part$data_source <- lab
    data_webcam_list <- append(data_webcam_list, list(data_webcam_part))
    
    data_demo_part <- read.csv2(file.path(DIR_NAME_DATA_MERGE, lab, FILE_NAME_DEMO))
    data_demo_part$data_source <- lab
    data_demo_list <- append(data_demo_list, list(data_demo_part))
      
    manual_exclusions_part <- read.csv2(file.path(DIR_NAME_DATA_MERGE, lab, FILE_NAME_EXCLUSION_MANUAL))
    manual_exclusions_part$data_source <- lab
    manual_exclusions_list <- append(manual_exclusions_list, list(manual_exclusions_part))
    
    sampling_rate_data_part <- read.csv(file.path(file.path(DIR_NAME_DATA_MERGE, lab, FILE_NAME_SAMPLING_RATE)))
    sampling_rate_data_part$data_source <- lab
    sampling_rate_data_list <- append(sampling_rate_data_list, list(sampling_rate_data_part))
    
    if (file.exists(file.path(DIR_NAME_DATA_MERGE, lab, FILE_NAME_EXCLUSION_AFTERPRE))){
      posterior_exclusions <- c(posterior_exclusions, readLines(file.path(DIR_NAME_DATA_MERGE, lab, FILE_NAME_EXCLUSION_AFTERPRE)))
    }
    
    f_exclusion_sampling <- file.path(DIR_NAME_DATA_MERGE, lab, FILE_NAME_EXCLUSION_SAMPLING)
    if (file.exists(f_exclusion_sampling)){
      
      samplingrate_exclusions_string <- readChar(f_exclusion_sampling, file.info(f_exclusion_sampling)$size)
      
      if(samplingrate_exclusions_string != "[]"){
        samplingrate_exclusions_string <- substr(samplingrate_exclusions_string, 3, nchar(samplingrate_exclusions_string)-2)
        
        if(grepl("', '", samplingrate_exclusions_string, fixed=T)){
          samplingrate_exclusions <- append(samplingrate_exclusions, unlist(strsplit(samplingrate_exclusions_string, "', '", fixed=T)))
        }else{
          samplingrate_exclusions <- append(samplingrate_exclusions, samplingrate_exclusions_string)
        }
      }
    }
  }
  
  rm(data_webcam_part, data_demo_part, manual_exclusions_part, samplingrate_exclusions_string, sampling_rate_data_part)
  
  data_webcam <- do.call("rbind", data_webcam_list)
  data_webcam <- data_webcam[ , !(names(data_webcam) %in% c("X"))]
  data_demo <- do.call("rbind", data_demo_list)
  data_demo <- as.data.frame(unclass(data_demo), stringsAsFactors = TRUE)
  manual_exclusions <- do.call("rbind", manual_exclusions_list)
  sampling_rate_data <- do.call("rbind", sampling_rate_data_list)
  sampling_rate_data <- sampling_rate_data[ , !(names(sampling_rate_data) %in% c("X"))]

  
  if(!dir.exists(DIR_NAME_DATA_MERGE_OUTPUT)){
    dir.create(DIR_NAME_DATA_MERGE_OUTPUT)
  }
  
  write.csv(data_webcam, file.path(DIR_NAME_DATA_MERGE_OUTPUT, FILE_NAME_WEBCAM), row.names = FALSE)
  write.csv(data_demo, file.path(DIR_NAME_DATA_MERGE_OUTPUT, FILE_NAME_DEMO), row.names = FALSE)
  write.csv(manual_exclusions, file.path(DIR_NAME_DATA_MERGE_OUTPUT, FILE_NAME_EXCLUSION_MANUAL), row.names = FALSE)
  write.csv(sampling_rate_data, file.path(DIR_NAME_DATA_MERGE_OUTPUT, FILE_NAME_SAMPLING_RATE), row.names = FALSE)
  
  writeLines(posterior_exclusions,
              file.path(DIR_NAME_DATA_MERGE_OUTPUT, FILE_NAME_EXCLUSION_AFTERPRE))
  
  samplingrate_exclusions <- samplingrate_exclusions[samplingrate_exclusions != ""]
  
  samplingrate_exclusions_out <- paste(samplingrate_exclusions, collapse="', '")
  writeLines(paste("['", samplingrate_exclusions_out, "']", sep=""),
              file.path(DIR_NAME_DATA_MERGE_OUTPUT, FILE_NAME_EXCLUSION_SAMPLING))
  
}else{ # singular data source
  data_webcam <- read.csv(FILE_NAME_WEBCAM, stringsAsFactors = TRUE)
  data_webcam <- data_webcam[ , !(names(data_webcam) %in% c("X"))]
  
  data_demo <- read.csv2(FILE_NAME_DEMO, stringsAsFactors = TRUE, sep = ",")
  
  manual_exclusions <- read.csv2(FILE_NAME_EXCLUSION_MANUAL, stringsAsFactors = TRUE)
  
  sampling_rate_data <- read.csv(FILE_NAME_SAMPLING_RATE, stringsAsFactors = TRUE)
    
  posterior_exclusions <- c()
  if (file.exists(FILE_NAME_EXCLUSION_AFTERPRE)){
    posterior_exclusions <- readLines(FILE_NAME_EXCLUSION_AFTERPRE)
  }
  
  samplingrate_exclusions <- c()
  if (file.exists(FILE_NAME_EXCLUSION_SAMPLING)){
    
    samplingrate_exclusions_string <- readChar(FILE_NAME_EXCLUSION_SAMPLING, file.info(FILE_NAME_EXCLUSION_SAMPLING)$size)
    
    if(samplingrate_exclusions_string != "[]"){
      samplingrate_exclusions_string <- substr(samplingrate_exclusions_string, 3, nchar(samplingrate_exclusions_string)-2)
    
      if(grepl("', '", samplingrate_exclusions_string, fixed=T)){
        samplingrate_exclusions <- unlist(strsplit(samplingrate_exclusions_string, "', '", fixed=T))
      }else{
        samplingrate_exclusions <- c(samplingrate_exclusions_string)
      }
    } 

    rm(samplingrate_exclusions_string)
  }
}


```

# Preprocessing

```{r}
# ManyWebcams data

## remove trial identifier from data
data_webcam$subid <- sub("_[^_]+$", "", data_webcam$subid)
sampling_rate_data$subid <- sub("_[^_]+$", "", sampling_rate_data$subid)

## exclude toddler from the final data set after relative data was already calculated
data_webcam <- filter(data_webcam, !subid %in% posterior_exclusions)

## long to wide format
data_webcam$condition <- NULL
data_webcam$row <- NULL
data_webcam <- data_webcam %>% spread(aoi, freq)

## calculate relative looking score for target vs. distractor only
data_webcam$percorr <- round(data_webcam$target/(data_webcam$target+data_webcam$distractor), digits = 2)
data_webcam$distractor <- NULL
data_webcam$target <- NULL
data_webcam$none <- NULL

## remove participants that only looked at "none", as they have no valid looking score (TODO How to report, only MPI_006 so far?)
data_webcam <- data_webcam[!is.nan(data_webcam$percorr),]

# sampling rate
sampling_rate_data <- filter(sampling_rate_data, subid %in% data_webcam$subid)
sampling_rate_mean <- sampling_rate_data %>%
  group_by(subid) %>%
  summarize(sampling_rate = mean(sampling_rate))

sum(sampling_rate_mean$subid %in% data_webcam$subid == F) # should be 0, otherwise check for mistakes
data_webcam$sampling_rate <- sampling_rate_mean$sampling_rate 

# MB2-Pilot data

# because some labs used exactly the same subid
data_pilot$subid_new <- paste0(data_pilot$subid, "_", data_pilot$lab)

data_inlab <- data_pilot %>%
  filter(experiment_num=="pilot_1a" & t >= -4000 & t <= 0) %>%
  group_by(subid_new) %>%
  summarize(lookingscore = sum(grepl("target", aoi, fixed = TRUE), na.rm = TRUE) / sum(grepl("target", aoi, fixed = TRUE) | grepl( "distractor", aoi, fixed = TRUE)),
            age_days = first(age),
            lab = first(lab))

# prepare for merging
data_inlab <- data_inlab %>%
  mutate(subid = subid_new,
         percorr = lookingscore) %>%
  select(subid, lab, percorr, age_days) 

data_webcam <- data_webcam %>%
  rename(lab = data_source)
  
## prepare inlab data for merging
data_inlab$percorr <- round(data_inlab$percorr, digits = 2)

# add method (webbased = 0, in-lab = 1)
data_webcam$method <- "webcam"
data_inlab$method <- "inlab"
```

# Descriptives for ManyWebcams

```{r}
# change data_demo$subid so that they match
data_webcamDemo_temp <- merge(data_webcam, data_demo, by = c("subid"))
# which don't match
x = which(data_webcam$subid %in% data_webcamDemo_temp$subid == FALSE)
x_subid = data_webcam$subid[x]

rm(data_webcamDemo_temp)

# add info about gender and age to data
data_webcamDemo <- merge(data_webcam, data_demo, by = c("subid"), all = TRUE)


data_webcamDemo <- data_webcamDemo %>%
  filter(!is.na(percorr))

data_webcamDemo$age <- data_webcamDemo$age_days
data_webcamDemo$gender <- data_webcamDemo$participant_gender
data_webcamDemo$country <- data_webcamDemo$residence_country

data_webcamDemo <- subset(data_webcamDemo, select = c(subid, age,  gender, country, percorr))

# calculate descriptives
## age in months
data_webcam$age_days <- data_webcamDemo$age
data_webcamDemo$age_months <- (data_webcamDemo$age/30.5)
data_webcamDemo$z_age <- scale(data_webcamDemo$age_months, scale = TRUE)

## Get number of girls and boys
sum(data_webcamDemo$gender == "girl")
sum(data_webcamDemo$gender == "boy") 

## Get residence country
summary(data_webcamDemo$country)

## Get age range
mean(data_webcamDemo$age, na.rm = TRUE)
sd(data_webcamDemo$age, na.rm = TRUE)
summary(data_webcamDemo$age)
```

```{r}
# because we don't have sampling data for the in lab testing
data_webcam_temp <- data_webcam %>%
  select(-sampling_rate)

# merge webbased & inlab
data <- rbind(data_webcam_temp, data_inlab)
rm(data_webcam_temp)

# change labs names
data[data$lab == "goettingen_babylab", "lab"] <- "GAUG"
data[data$lab == "lmu_babylab", "lab"] <- "LMU"
data[data$lab == "ubc_infantlab", "lab"] <- "UBC"
data[data$lab == "trento_babylab", "lab"] <- "BLT"

# z-stand. of age
data$age_months <- (data$age_days/30.5)
data$z_age <- scale(data$age_days, scale = TRUE)
data_webcam$z_age <- scale(data_webcam$age_days, scale = TRUE)
```

# Inferential Statistics
H1: Children will engage in goal-based action predictions, indicated by above-chance looking towards the location that matches the outcome of an agent´s action goal
```{r, warning=FALSE}
t.test(data_webcam$percorr, mu = 0.5, alternative = "two.sided")
cohen.d(data_webcam$percorr ~ 1, mu = 0.5)
```

H2: Does the eye-tracking method have an effect on the measured proportional looking score? No strong directional hypothesis. (Method will be contrast coded (web-based = 0, lab-based = 1))
```{r, warning=FALSE}
# Error: number of levels of each grouping factor must be < number of observations (problems: subid)
# definition of new ID

#data$ID <- NA
#data$ID <- c(1:95)
#h2 <- lmer(percorr ~ method + (1|ID), data=data) #same error --> reason: only one observation per participant (no random slope)

data <- data%>%
  mutate(method = factor(method, levels = c("webcam", "inlab")))

h2 <- lm(percorr ~ method, data)

anova(h2)
summary(h2)
plot(allEffects(h2))
```
## random intercept for lab

```{r}
h2.rand.model <- lmer(percorr ~ 0+ offset(0.5) + method  + (1|lab), data = data)
summary(h2.rand.model)
```
## beta distribution

```{r}
# modifying the data because beta regression requires values larger than 0 and smaller than 1
mdata <- data%>%
  mutate(percorr = ifelse(percorr == 0, 0.000001, percorr),
         percorr = ifelse(percorr == 1, 0.999999, percorr))%>%
  mutate(method = factor(method, levels = c("webcam", "inlab")))

h2.rand.model.beta <- glmmTMB(percorr ~ offset(0.5) + method + (1|lab), data = mdata, family = beta_family())
summary(h2.rand.model.beta)
```

H3: Lower proportion of children who contribute usable data will be lower in webcam-based as compared to lab-based eye-tracking.

```{r}

nTestedWebcam = nrow(data_demo) - length(posterior_exclusions)
nIncludedWebcam = nrow(data_webcam)

exclusion.table <- matrix(c(nIncludedWebcam, nrow(data_inlab), 91, 8), nrow = 2, ncol = 2, byrow = T)
rownames(exclusion.table) <- c('included', 'excluded')
colnames(exclusion.table) <- c('online', 'in-lab')

exclusion.table


h3 <- chisq.test(exclusion.table)
h3 
# p-value sign., both variables are not independent
h3$expected
```


# preparation for plot
```{r}
# create values for error bars
mean(data_webcam$percorr) 
mean(data_inlab$percorr)

dataErrorbar <- group_by(data, method) %>%
  summarise(mean = mean(percorr), se = sd(percorr)/sqrt(n()), upper = mean + se, lower = mean - se)

# create values for 95% CIs
require(gmodels)
plot.data <- as.data.frame(round(ci(h2),3))
plot.data$method <- c("webcam", "inlab")

plot.data[plot.data$method == "inlab", "Estimate"] <- plot.data[plot.data$method == "webcam", "Estimate"] + plot.data[plot.data$method == "inlab", "Estimate"]
plot.data[plot.data$method == "inlab", "CI lower"] <- plot.data[plot.data$method == "webcam", "CI lower"] + plot.data[plot.data$method == "inlab", "CI lower"]
plot.data[plot.data$method == "inlab", "CI upper"] <- plot.data[plot.data$method == "webcam", "CI upper"] + plot.data[plot.data$method == "inlab", "CI upper"]
```

# plot
```{r, warning=FALSE}
# plot the data
H2.plot <- ggplot(data, aes(x = method, y = percorr, colour = method, fill = method)) +
  geom_violin(alpha = 0.5, width = 1, position = position_dodge(width = 0.9), show.legend = FALSE ) +
  geom_jitter(aes(colour = method),  size = 2, alpha = 0.6, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0, dodge.width = 0.9), show.legend = FALSE) +
 geom_point(aes(x = method, y = mean), dataErrorbar, inherit.aes = FALSE, size = 2, color = "black") +
  # here are now CIs instead of SE
  geom_errorbar(aes(x = method, ymax = `CI upper`, ymin = `CI lower`), plot.data, inherit.aes = FALSE, 
             stat = "identity", width = 0.05, color = "black") +
  ylim(0, 1) +
  scale_colour_manual(values=c("skyblue", "red"))  +
  scale_fill_manual(values=c("skyblue",  "red")) +
  geom_hline(yintercept=0.5, linetype="dashed", color = "black") +
  labs(title="Proportion Looking Score per method (95% CIs)", x="Method", y="Proportion Looking Score (target/target+distractor)") +
  scale_x_discrete(labels=c("lab" = "in-lab", "online" = "online")) +
  theme_classic()

print(H2.plot)
```

# Bayesian analysis

Bayesian version of the model because it gives us more flexibility with respect to the response distribution (`lm` simply assumes normal distribution but ours is truncated given that we work with proportions).

```{r}
#visualize response distribution
# distribution is truncated by 0 and 1 and skewed - Beta distribution might be suited to model this

ggplot(data, aes(x = percorr))+
  geom_histogram(fill = NA, col = "black")+
  theme_bw()
```


```{r}
library(brms)
library(tidybayes)

# modifying the data because beta regression requires values larger than 0 and smaller than 1
mdata <- data%>%
  mutate(percorr = ifelse(percorr == 0, 0.000001, percorr),
         percorr = ifelse(percorr == 1, 0.999999, percorr))

bm1 <- brm(percorr ~ method,
           data = mdata,
           family = Beta(),
           iter = 2000, warmup = 1000, cores = 4, chains = 4
           )

# check how well the model is able to reproduce the data
pp_check(bm1)

# model summary
bm1
```


```{r}
# create a new dataset with posterior predictions for the mean for each setting
post <-mdata%>%
  distinct(method)%>%
  add_epred_draws(bm1)

# plot the predicted model average (with 95% CrI) and the data
ggplot(post, aes(x = .epred, y = method, fill = method))+
  geom_vline(xintercept = 0.5, lty = 3, alpha = .5)+
  geom_point(data = mdata, aes(x = percorr, y = method, col = method),pch = 1,alpha = .5)+
  stat_halfeye(alpha = .75, .width = c(0.66, 0.95))+
  scale_colour_manual(values=c("skyblue", "red"))  +
  scale_fill_manual(values=c("skyblue",  "red")) +
  labs(x = "Proportion Looking Score (target/target+distractor)", y = "Method")+
  theme_bw()
```

# Exploratory analyses

## adding age as a predictor

1. To investigate potential age effects, we will include age as a predictor in an exploratory model.

```{r}
ex.age.model <- lm(percorr ~ z_age, data = data_webcamDemo)

summary(ex.age.model)

ex.age.rand.model <- lmer(percorr ~ method + z_age + (1|lab), data = data)
summary(ex.age.rand.model)

```

## recordings sampling rate

3. We plan to analyze the effect of the recording’s sampling rate on the proportional looking score in an additional model.

```{r}
mean(data_webcam$sampling_rate)
sd(data_webcam$sampling_rate)
summary(data_webcam$sampling_rate)

ex.sampling.model <- lmer(percorr ~ z_age + sampling_rate + (1|lab), data = data_webcam)

summary(ex.sampling.model)
```

Mean sampling rates (per subject) had no effect on the proportional looking score.

