---
title: "ManyWebcams: web-based eye-tracking"
author: "CDS preliminary analysis"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: yes
editor_options: 
  chunk_output_type: inline
---

```{r libraries, results='hide', message=FALSE, warning=FALSE}
library(readr)
library(knitr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(psych)
library(effsize)
library(Matrix)
library(lme4)
library(lmerTest)
library(carData)
library(effects)
library(stringr)

knitr::opts_chunk$set(cache = TRUE, warn = FALSE, message = FALSE)
options(scipen=999)
```

# Intro
Analysis script for the ManyWebcams web-based eye-tracking study. Script for preliminary results to present at CDS 2022

# working directory
```{r}
#setwd("G:/ManyWebcams/Analyses/CDS_prelim_analysis")
```

# File reading
```{r}
data_webcam <- read.csv("relative_data.csv", stringsAsFactors = TRUE)
data_pilot <- read.csv("pilot_data.csv", stringsAsFactors = TRUE)
data_demo <- read.csv2("demographic_data.csv", stringsAsFactors = TRUE)
```

# Preprocessing
```{r}
# ManyWebcams data
## exclude one toddler, because preterm
data_webcam <- filter(data_webcam, !subid %in% c("LMU_746_A"))

## long to wide format
data_webcam$condition <- NULL
data_webcam$row <- NULL
data_webcam <- data_webcam %>% spread(aoi, freq)

## calculate relative looking score for target vs. distractor only
data_webcam$percorr <- round(data_webcam$target/(data_webcam$target+data_webcam$distractor), digits = 2)
data_webcam$distractor <- NULL
data_webcam$target <- NULL
data_webcam$none <- NULL

# MB2-Pilot data
data_inlab <- data_pilot %>%
  filter(experiment_num=="pilot_1a" & t >= -4000 & t <= 0) %>%
  group_by(subid) %>%
  summarize(
    lookingscore = sum(grepl("target", aoi, fixed = TRUE), na.rm = TRUE) / sum(grepl("target", aoi, fixed = TRUE) | grepl( "distractor", aoi, fixed = TRUE))
  )

## prepare inlab data for merging
colnames(data_inlab)[colnames(data_inlab) == 'lookingscore'] <- 'percorr'
data_inlab$percorr <- round(data_inlab$percorr, digits = 2)

# add method (webbased = 0, in-lab = 1)
data_webcam$method <- "webcam"
data_inlab$method <- "inlab"

# merge webbased & inlab
data <- rbind(data_webcam, data_inlab)
```

# Descriptives for ManyWebcams
For the children sample, a total of 50 participants were tested. From these, 15 had to be excluded due to technical errors with the eye tracking procedure (e.g., webcam could not be accessed, data upload) and 3 due to low sampling rate. From the remaining 32 participants, 2 had to be excluded due to interference (e.g., caregiver interfered with procedure or child was inattentive) during the procedure. These trials were identified by visual inspection. The final sample of included children is N = 30.
```{r}
# add info about gender and age to data
data_webcamDemo <- merge(data_webcam, data_demo, by = c("subid"))

data_webcamDemo$age <- data_webcamDemo$age_days
data_webcamDemo$gender <- data_webcamDemo$participant_gender
data_webcamDemo$country <- data_webcamDemo$residence_country

data_webcamDemo <- subset(data_webcamDemo, select = c(subid, age,  gender, country, percorr))

# calculate descriptives
## age in months
data_webcamDemo$age <- (data_webcamDemo$age/30.5)

## Get number of girls and boys
sum(data_webcamDemo$gender == "girl") #17
sum(data_webcamDemo$gender == "boy") #13

## Get residence country
summary(data_webcamDemo$country) # Germany: 19, UK: 8, Austria: 3

## Get age range
mean(data_webcamDemo$age) #22.27 months
sd(data_webcamDemo$age) #2.400 months
summary(data_webcamDemo$age) # min.:18.69 months, max.: 26.92 months

## Descriptives for ManyBabies2-Pilot
## n=65 (37 girls; mean age = 23.14 months; range 18.25 ro 26.84 months)
```

# Inferential Statistics
H1: Children will engage in goal-based action predictions, indicated by above-chance looking towards the loaction that matches the outcome of an agent´s action goal
```{r, warning=FALSE}
t.test(data_webcam$percorr, mu = 0.5, alternative = "two.sided")
cohen.d(data_webcam$percorr ~ 1, mu = 0.5)
```

H2: Does the eye-tracking method have an effect on the measured proportional looking score? No strong directional hypothesis. (Method will be contrast coded (web-based = 0, lab-based = 1))
```{r, warning=FALSE}
# Error: number of levels of each grouping factor must be < number of observations (problems: subid)
# definition of new ID

#data$ID <- NA
#data$ID <- c(1:95)
#h2 <- lmer(percorr ~ method + (1|ID), data=data) #same error --> reason: only one observation per participant (no random slope)

h2 <- lm(percorr ~ method, data)

anova(h2)
summary(h2)
plot(allEffects(h2))

str(data$subid)
```

H3: Lower proportion of children who contribute usable data will be lower in webcam-based as compared to lab-based eye-tracking.

ManyWebcams: 60% included; exclusion due to technical problems (n= 15), low sampling rate (n= 3), manual exclusion: interference (n = 2)
ManyBabies2 Pilot: 89.04% included; exclusion due to early termination (n = 6), technical problems (n= 2)

```{r}
exclusion.table <- matrix(c(30, 65, 20, 8), nrow = 2, ncol = 2, byrow = T)
rownames(exclusion.table) <- c('included', 'excluded')
colnames(exclusion.table) <- c('online', 'in-lab')

exclusion.table

h3 <- chisq.test(exclusion.table)
h3 
# p-value sign., both variables are not independent
h3$expected
```


# preparation for plot
```{r}
# create values for error bars
mean(data_webcam$percorr) # 0.5898
mean(data_inlab$percorr)  # 0.7345

dataErrorbar <- group_by(data, method) %>%
  summarise(mean = mean(percorr), se = sd(percorr)/sqrt(n()), upper = mean + se, lower = mean - se)
```

# plot
```{r, warning=FALSE}
# plot the data
H2.plot <- ggplot(data, aes(x = method, y = percorr, colour = method, fill = method)) +
  geom_violin(alpha = 0.5, width = 1, position = position_dodge(width = 0.9), show.legend = FALSE ) +
  geom_jitter(aes(colour = method),  size = 2, alpha = 0.6, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0, dodge.width = 0.9), show.legend = FALSE) +
 geom_point(aes(x = method, y = mean), dataErrorbar, inherit.aes = FALSE, size = 2, color = "black") +
  geom_errorbar(aes(x = method, ymax = upper, ymin = lower), dataErrorbar, inherit.aes = FALSE, 
             stat = "identity", width = 0.05, color = "black") +
  ylim(0, 1) +
  scale_colour_manual(values=c("skyblue", "red"))  +
  scale_fill_manual(values=c("skyblue",  "red")) +
  geom_hline(yintercept=0.5, linetype="dashed", color = "black") +
  labs(title="Proportion Looking Score per method", x="Method", y="Proportion Looking Score (target/target+distractor)") +
  scale_x_discrete(labels=c("lab" = "in-lab", "online" = "online")) +
  theme_classic()

print(H2.plot)
```
